### 2016-06-15 testbed_nfs.tex
Started on nfs setup

\label{sec:testbed_nfs}

\subsection{NFS}

% Define variable]
\begin{lstlisting}[]
$ export IMAGE="2016-05-27-raspbian-jessie-lite"
$ export WORKDIR="${HOME}/Raspbian"
$ export DEV="/dev/mmcblk0"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

% Check out the image   % The dollar hack was to fix vim syntax
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u sectors -l ${DEV}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}