### 2016-07-07 overview.tex
Document structure changes and new appendix changes

\label{sec:overview}

A sketch of the testbed is depicted in Fig.~\ref{fig:testbed_setup}.
The testbed consists of up to 100 \ac{Raspi}s of different models.
More specifically, in our design we consider: Raspberry Pi 1 model B rev.~2,
Raspberry Pi 2 model B V1.1 and Raspberry Pi 3 model B V1.2.
All \ac{Raspi}s are each equipped with a memory card, a wired and wireless
network interface and a power supply. All the \ac{Raspi} are connected to
a common \ac{LAN} that provides internal and external connectivity. Without
loss of generality, in our case they are connected to a university network
using their wired Ethernet interface that is named \texttt{eth0} according
to the legacy naming convention of Ethernet interfaces in
Linux~\cite{PredictableNetworkInterfaceNames}. We consider the
university network since our testbed is used by students and
professors to perform measurements and experimentation of controlled and
reproducible scenarios as part of academic research. Nevertheless, the
testbed description and procedures for setting it up are not restricted
to the academic scenario. All \ac{Raspi}s will be configured to run
a \ac{SSH} daemon for easy remote access within the university network.
We requested the university \ac{IT} department to configure the university
\ac{DHCP} server to assign each \ac{Raspi} a static \ac{IP}. This
eliminates the requirements of monitors and keyboards with the \ac{Raspi}s
for non-graphical applications. Finally, our design aims to configure all
\ac{Raspi}s identically from a customized bootable image in their
respective memory cards, while still allowing the end-users to store
files locally in each of the \ac{Raspi}s.

\begin{figure}[ht!]
\centering
\includegraphics[width=0.45\textwidth]{images/testbed_setup3.eps}
\caption{Testbed setup}
\label{fig:testbed_setup}
\end{figure}

%This significantly eases the management of the \ac{Raspi}s
%as \ac{SSH} eliminates the need for monitors and keyboards for the \ac{Raspi}s.

In Fig.~\ref{fig:testbed_setup} and what follows, we will refer as the
testbed administrator to the person(s) in charge of setting up and
configuring the testbed with administrator privileges from the \ac{OS}
point of view. The setting and configuration procedure commands are
performed by the testbed administrator in a regular \ac{PC} running a
Linux distribution. Although in principle the administrator
Linux distribution is not a restriction, we present our procedure in a
Debian-based Linux distribution. Our basic design considers to create
a customized image to later store it in a memory card. Thus, based on a
regular Raspbian Linux image, we explain how to resize and customize it to
include desired packages for performing routinary tasks in all the
\ac{Raspi}s. As another design consideration, when working on a testbed
with many devices, it may not always be desired to make persistent changes
on them. One scenario is when different users are interested in
running experiments on a rebooted testbed. Thus, we later present how
to utilize stacked filesystems to enable both persistent and
temporary storage to have this capability. Its purpose is to
remove non-desired data after a reboot while keeping the original
customized image structure. This step of the procedure is optional if
the testbed administrator decides to keep only persistent changes regardless
of the testbed use. Finally, an advanced step in our design considers to
fetch the \ac{Raspi}s root filesystem from a \ac{HTTP} server. This enables the
testbed administrator to put the data files in one location. The
\ac{Raspi}s will then get these data files after a reboot. We consider
that this step is particularly useful when there is a large amount
of devices to be configured and the setup process needs to be automated.
However, since a more complex scenario that might be required only in special
cases, we put this step in Section~\ref{sec:http} in the Appendix.