### 2016-07-15 appendix.tex
Preparing http script

\section{HTTP Image Fetch Setup (Optional)}
\label{sec:http}

\input{testbed_http}

\section{Alternative Commands for Outdated Packages}
This section describes alternative commands in case the testbed administrator
is not able to update old packages in its Linux distribution for performing
the commands, particularly the ones regarding the \texttt{util-linux} package
or if a command just fails.

\subsection{Losetup for Loopback Devices}
\label{sec:alternative_losetup}

In case the \texttt{losetup -P} is not available given that a version
higher or equal than 2.21 for \texttt{util-linux} is neither not available,
a possible option is to manually set all the loopback devices used during the
whole procedure with \texttt{mknod}. To do this, the alternative commands are:

\begin{lstlisting}[]
$ export DEV="/dev/loop0"
# mknod ${DEV}p1 b 7 1
# mknod ${DEV}p2 b 7 2
# losetup -o $((8192*512)) --sizelimit $(( (137215-8192+1)*512) )) ${DEV}p1 ${IMAGE}.img
# losetup -o $((137216*512)) --sizelimit $(( (4806655-137216+1)*512) )) ${DEV}p2 ${IMAGE}.img
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

However, there will be a few differences by using the above code as an
alternative to the \texttt{losetup -P} case. First, the output from
\texttt{lsblk} will look different:

\begin{lstlisting}[]
# lsblsk
NAME   MAJ:MIN RM   SIZE RO TYPE MOUNTPOINT
...
loop1    7:1    0    67M  0 loop
loop2    7:2    0   2.2G  0 loop
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Second, detaching the loop devices will also be different:

\begin{lstlisting}[]
# losetup -d ${DEV}p{1,2}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Working with these loopback devices will be transparent for the remainding
procedure.

Also, when creating the persistent home directory partition in
Section~\ref{sec:persistent_directories}, for recreating the loopback devices
the commands are:

\begin{lstlisting}[]
$ export DEV="/dev/loop0"
# mknod ${DEV}p2 b 7 2
# mknod ${DEV}p3 b 7 3
# losetup -o $((137216*512)) --sizelimit $(( (4806655-137216+1)*512) ) ${DEV}p2 ${IMAGE}.img
# losetup -o $((4806656*512)) --sizelimit $(( (6903807-4806656+1)*512)) ${DEV}p3 ${IMAGE}.img
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Finally, for detaching in this case the command is:
\begin{lstlisting}[]
# losetup -d ${DEV}p{2,3}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Image Chroot with Proot}
\label{sec:chroot}

In some cases, \texttt{chroot} may not work properly. In this case, an
alternative can be \texttt{proot}. It might be required to install it
before changing the root with \texttt{apt-get install proot}. Then, it
should be possible for the testbed administrator to run the following:

% CHROOT to OS image
\begin{lstlisting}[]
# proot -q qemu-arm-static -S ${ROOTDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Umount after Image Chroot}
\label{sec:umount}

In case the \texttt{umount --recursive} is not available given that a version
higher or equal than 2.22 for \texttt{util-linux} is neither not available,
a possible option is to manually umount all the partitions used during the
\texttt{chroot} environment. The directories must be unmounted in the reverse
order as when they were mounted. These commands are:

\begin{lstlisting}[]
# umount ${ROOTDIR}/dev/pts
# umount ${ROOTDIR}/dev
# umount ${ROOTDIR}/sys
# umount ${ROOTDIR}/proc
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Paramiko Package Update}
\label{sec:paramiko}
In case Paramiko version 1.15.1 or higher is not installed, we may observe
a key exchange algorithm error. In this case, it might be
necessary to update the Paramiko package and verify this by running:

\begin{lstlisting}[]
# pip install --upgrade paramiko
...
Sucesfully installed cffi-1.7.0 cryptography-1.4 paramiko-2.0.1
# pip show paramiko
...
Name: paramiko
Version: 2.0.1
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

This should allow to make \ac{SSH} connections with Fabric by using the proper
key exchange algorithms through Open SSH.