### 2016-08-25 reviewer_reply.txt
added reviewer replies

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 1:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for you valuable feedback and detailed comments.
They were very useful. We have entirely rewritten a few of the sentences
indicated, but in most cases made the corrections as suggested.

We like the suggested citations. The first two citations are now used
to improve our introduction, but we do not see how to fit in the last
citation "Data Exfiltration from Internet of Things Devices: iOS
Devices as Case Studies" which we think is related to IoT, but not so
much Raspberry Pi, testbeds, or network coding.

To address the applications for a Raspberry Pi testbed, we have added a few
other citations of practical usecases of the Raspberry Pi. Furthermore,
we will ask the editors to have a citation to another MDPI journal paper of
ours if it is being accepted. This paper evaluates the speed and energy
consumption of three network coding schemes in Raspberry Pi devices from our
testbed.

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 2:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you for you valuable feedback

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 3:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,
Thank you very much for your valuable feedback and suggested citations.
The additional citations has significantly helped to better address the
usecases for a Raspberry Pi testbed. We think that the
papers are great, but the core idea behind the setups in the first four
seem quite similar. We suggest to add suggested paper 2, 3, 5, and a few
papers indicated by another reviewer.