### 2016-08-26 reviewer_reply.txt
Final revised version. Few corrections, typos and removing an extrablank page

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 1:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,

Thank you very much for you valuable feedback and detailed comments.
They have been very useful. To address your comments, we list
the ones that require our modification below and provide our answer.

- "To improve the article and considering that already others authors have faced problems similar even if different, the reviewer suggests to enrich the bibliography with others articles. For instance, in the introduction could be possible to cite these:

1) ...'An Intelligent and High Efficiency Street Lighting System Isle Based on...' ...
2) ...'A Fully Controlled Street Lighting System Isle Based on...' ...
3)  ...'A New Acquisition and Imaging System for Environmental...' ...
4) ...'A New Remote and Automated Control System for the Vineyard...'
5) ...'FingerScanner: Embedding a Fingerprint Scanner in a Raspberry Pi' "

We consider relevant the suggested references. Still, we include different use
cases and journal versions of articles with similar core ideas behind. Thus, after
reading them, references 2-3) and 5) are now used as references 11-13)
in the new revised manuscript to improve our introduction. Besides these, in the
revised manuscript, references 9-10) have been included to enhance
the quality of it.

- "As minor revision please check and correct the following rows: 62, 171, 546, 705."

We have corrected all the typos by proof-reading the manuscript.
Also, we have entirely rewritten a few of the sentences
indicated as misunderstanding, but in most cases made the corrections as suggested.

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 2:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,

Thank you for your time with the revision.

- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - 
Reviewer 3:
- - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

Dear reviewer,

Thank you very much for your valuable feedback and suggested citations.
To address your comments, we list the ones that require our modification
below and provide our answer. For simplicity, we list them by type
of comment

- Comments of the form: "The suggestions by :

...
Page XX,
Line YY
Line ZZ
...
"

We have corrected all the typos by proof-reading the manuscript.
Also, we have entirely rewritten a few of the sentences
indicated as misunderstanding, but in most cases made the corrections as suggested.

- Comment of the form: "Suggest to add:

1) Cox, S. ...'Iridis-pi: a low-cost, compact demonstration cluster'...
2) Raspberry Pi Foundation ...'The Making of Pi,'...
3) D'Orazio C ...'Data Exfiltration from Internet of Things Devices: iOS Devices as Case Studies'...

"

These citations have helped to better address the
prior work for a Raspberry Pi testbed. We think 1) and 2) are relevant
references and are used in the new revised manuscript as references 9-10).
Nevertheless, but we do not see how to fit the reference
"Data Exfiltration from Internet of Things Devices: iOS
Devices as Case Studies" which we think is related to IoT, but not very much
to Raspberry Pi, testbeds or network coding. Besides these, references 11-13)
in the new revised manuscript, have been included to enhance the quality of it.