### 2016-09-18 appendix.tex
changed toolchain link

% \section{HTTP Image Fetch Setup (Optional)}
% \label{sec:http}

% \input{testbed_http}

\section{Alternative Commands for Outdated Packages}
This section describes alternative commands in case the testbed administrator
is not able to update old packages on his / her Linux distribution for performing
the commands, particularly the ones regarding the \texttt{util-linux} package
or if a command just fails.

\subsection{Losetup for Loopback Devices}
\label{sec:alternative_losetup}

In case the \texttt{losetup -P} command shows the \texttt{invalid option}
message, an alternative is to manually set all the loopback devices
used during the whole procedure with \texttt{mknod}. To do this, the
alternative commands are:

\begin{lstlisting}[]
$ export DEV="/dev/loop0"
# mknod ${DEV}p1 b 7 1
# mknod ${DEV}p2 b 7 2
# losetup -o $((8192*512)) --sizelimit $(( (137215-8192+1)*512) )) ${DEV}p1 ${IMAGE}.img
# losetup -o $((137216*512)) --sizelimit $(( (4806655-137216+1)*512) )) ${DEV}p2 ${IMAGE}.img
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

However, there will be a few differences by using the above code as an
alternative to the \texttt{losetup -P} case. First, the output from
\texttt{lsblk} will look different:

\begin{lstlisting}[]
# lsblsk
NAME   MAJ:MIN RM   SIZE RO TYPE MOUNTPOINT
...
loop1    7:1    0    67M  0 loop
loop2    7:2    0   2.2G  0 loop
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Second, detaching the loop devices will also be different:

\begin{lstlisting}[]
# losetup -d ${DEV}p{1,2}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Working with these loopback devices will be transparent for the remainding
procedure.

Also, when creating the persistent home directory partition in
Section~\ref{sec:non_persistent_directories}, for recreating the loopback
devices the commands are:

\begin{lstlisting}[]
$ export DEV="/dev/loop0"
# mknod ${DEV}p2 b 7 2
# mknod ${DEV}p3 b 7 3
# losetup -o $((137216*512)) --sizelimit $(( (4806655-137216+1)*512) ) ${DEV}p2 ${IMAGE}.img
# losetup -o $((4806656*512)) --sizelimit $(( (6903807-4806656+1)*512)) ${DEV}p3 ${IMAGE}.img
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Finally, for detaching in this case the command is:
\begin{lstlisting}[]
# losetup -d ${DEV}p{2,3}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Image Chroot with Proot}
\label{sec:chroot}

In some cases, \texttt{chroot} may not work properly. In this case, an
alternative can be \texttt{proot}. It might be required to install it
before changing the root with \texttt{apt-get install proot}. Then, it
should be possible for the testbed administrator to run the following:

% CHROOT to OS image
\begin{lstlisting}[]
# proot -q qemu-arm-static -S ${ROOTDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Umount after Image Chroot}
\label{sec:umount}

%The recursive command (\texttt{----recursive}) is not valid for \texttt{umount}
%is not valid for \texttt{util-linux} version 2.22
%or later. An alternative option is to manually unmount all the partitions used
%during the \texttt{chroot} environment. The directories must be unmounted in
%the reverse order as when they were mounter. These commands are:

%The --recursive parameter is not part of umount until 
%\texttt{umount ----recursive} is not part of \texttt{util-linux} version 2.22

In case the \texttt{umount ----recursive} command shows the \texttt{invalid option}
message, an alternative is to manually unmount all the partitions used
during the \texttt{chroot} environment in the reverse order which they were mounted.
These commands are:
%In case the \texttt{umount ----recursive} command fails given that a version
%higher than 2.22 for \texttt{util-linux} is not available,
%a possible option is to manually umount all the partitions used during the
%\texttt{chroot} environment. The directories must be unmounted in the reverse
%order as when they were mounted. These commands are:

\begin{lstlisting}[]
# umount ${ROOTDIR}/dev/pts
# umount ${ROOTDIR}/dev
# umount ${ROOTDIR}/sys
# umount ${ROOTDIR}/proc
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

\subsection{Paramiko Package Update}
\label{sec:paramiko}
In case Paramiko version 1.15.1 or higher is not installed, we may observe
a key exchange algorithm error when trying to log in through \ac{SSH} to a
Raspi with Open SSH. In this case, it might be necessary to update the Paramiko
package that Fabric uses to remove this error by running:

\begin{lstlisting}[]
# pip install --upgrade paramiko
...
Successfully installed cffi-1.7.0 cryptography-1.4 paramiko-2.0.1
# pip show paramiko
...
Name: paramiko
Version: 2.0.1
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}